Local simple connectedness of boundaries of hyperbolic groups

### 2020-04-01 paper.tex
Finished the introduction and proof-read sections 1,2

\todo{Proof-read all}

@@

\begin{enumerate}
  \item Introduce boundaries. What are they, why are they important.
  \item Kapovich's question. Not much flexibility, not very many restrictions
    either.
  \item Partial answers.
\end{enumerate}

The goal of this paper is to prove the following theorem, which restricts the
local complexity of the boundary of a hyperbolic group in terms of its global
complexity.

@@

In Section~\ref{sec:spheres} we use the Rips complex of a group $G$ to build a
sequence of simplicial complexes $(K_n)_{n \geq 0}$ that give successively
better approximations to the boundary $\boundary G$. These simplicial complexes
are spheres in the Rips complex. There exist natural simplicial maps
$p^n_m\from K_n\to K_m$ for $n - m$ large enough, and there exist natural
continuous maps $p^\infty_n\from\boundary G \to K_n$ for all $n$.  The
remainder of Section~\ref{sec:spheres} will be spent proving quantitative
statements about these maps.

In Section~\ref{sec:horoball} we prove a compactness result: given a sequence
of vertices $v_n \in K_n$, there exist group elements $g_n$ and a subcomplex
$H$ of the Rips complex such that translation by $g_n$ maps a large
neighbourhood in $K_n$ of $v_n$ onto a large neighbourhood in $H$ of $g_n\cdot
v_n$.  We may further assume that $(g_n)$ converges to a point $\xi \in
\boundary G$, and then we think of $H$ as a horoball in $G$ asymptotic to
$\xi$. We then spend the remainder of Section~\ref{sec:horoball} describing
projection maps $\boundary G - \set\xi \to H$.

In Section~\ref{sec:double_dagger} we begin by recalling the so-called
double-dagger condition $\ddag$ of Bestvina and Mess. This condition is
guaranteed to hold if $\boundary G$ does not contain a cut point, and implies
that $\boundary G$ is locally path connected. We then modify the condition to
allow for easier generalisation, and introduce a new condition $\S$, which we
will see to be related the local simple connectedness of $\boundary G$. Using
the compactness results from Section~\ref{sec:horoball} we show that this
condition is satisfied whenever we have that $\boundary G - \set\xi$ is simply
connected for any $\xi$ in $\boundary G$.

Then, in Section~\ref{sec:local_simple_connectedness} we show that our new
condition $\S$ implies the local simple connectedness of $\boundary G$. We
begin by using $\S$ to construct a sequence of maps $(i^n_\infty)_n$ from the
2-skeleton of $K_n$ to $\boundary G$. These maps can be thought of as
approximate sections of $p^\infty_n$, in that $i^n_\infty\composed p^\infty_n$
converges uniformly to the identity map $\boundary G \to \boundary G$ as $n$
tends to $\infty$. By controlling the diameter of the image under $i^n_\infty$
of a simplex in $K_n$ we are able to prove the local simple connectedness of
$\boundary G$.

@@

Let $G$ be a hyperbolic group and let $S$ be a symmetric generating set for
$G$. Let $\dist(\cdot, \cdot)$ be the natural metric on the Cayley graph of $G$
with respect to $S$; on $G$ this metric agrees with the word metric.

Let $\delta \geq 0$ be such that the Cayley graph of $G$ with respect to $S$ is
$\delta$-hyperbolic, meaning that all geodesic triangles in the Cayley graph of
$G$ are $\delta$-thin. To reduce the growth of constants later on we in
fact assume that all \emph{ideal} triangles (with vertices in $G \union
\boundary G$) are $\delta$-hyperbolic; this can be guaranteed by multiplying
$\delta$ by $4$.

Fix an ordering on $S$. Let $S^\star$ be the free monoid on 
$S$; order $S^\star$ by the lex-least ordering: this ordering is defined by 
requiring that shorter words precede longer and that words of equal length are 
ordered lexicographically. Let $D = 10^6\delta + 10^6$.

\begin{definition}
  We denote by $K$ the \emph{Rips complex} of $G$; this is the simplicial
  complex with vertex set $G$ such that a finite set of elements of $G$ spans a
  simplex if and only if that set has diameter at most $D$ with respect to the
  word metric $d$.
\end{definition}

@@

\begin{remark}\label{rem:dist_vs_newdist}
  The extended metric $\newdist$ is not a metric, but this will not cause a
  problem.

  Notice that $\newdist$ agrees with $\dist$ on $G$, and for any $x$ and $y$ in
  $K$, if $x^\star$ and $y^\star$ are \emph{any} vertices of simplices
  containing $x$ and $y$ respectively, then
  \begin{equation*}
    \abs{\newdist(x,y) - \dist(x^\star, y^\star)} \leq 2D
  \end{equation*}
\end{remark}

@@

\begin{definition}
  Given points $x$, $y$ and $z$ in $K$, define the \emph{Gromov product
  $\gromov{x}{y}{z}$ of $x$, $y$ and $z$} as follows.
  \begin{equation*}
    \gromov{x}{y}{z} = \frac{1}{2}\left(\newdist(x,y) + \newdist(x,z) -
          \newdist(y,z)\right)
  \end{equation*}
\end{definition}

We now equip $K$ with a topology as follows.

@@

\begin{definition}
  For $n \geq m$ define a map $p^n_m \from S_n \to S_m$ as follows.
  Given an element $g$ of $S_n$ let $w_g \in S^\star$ be the lexicographically
  first representative of $g$; note that this word in $S^\star$ necessarily has
  length $n$.  Then define $p^n_m(g)$ to be the element of $G$ given by
  truncating $w_g$ to a word of length $m$.
\end{definition}

@@

\begin{proof}
  Let $\set{g_1, \dotsc, g_k}$ span a simplex in $K_n$.  Then for $g_i$ and
  $g_j$ in $\set{g_1, \dotsc, g_k}$, the distance from $p^n_m(g_i))$ to a
  geodesic from $g_i$ to $g_j$ is at least $n - m - D$, which is greater than
  $\delta$, so by hyperbolicity of a triangle with vertices $e$, $g_i$ and $g_j$,
  the distance from $p^n_m(g_i)$ to any geodesic from $e$ to $g_j$ is at most
  $\delta$. We may choose this geodesic to contain $p^n_m(g_j)$, and $\dist(e,
  p^n_m(g_i)) = \dist(e, p^n_m(g_j))$, so $\dist(p^n_m(g_i), p^n_m(g_j)) \leq
  2\delta \leq D$.  It follows that the diameter of $\set{p^n_m(g_1), \dotsc,
  p^n_m(g_k)}$ is at most $D$ and therefore the set spans a simplex in $K_m$.
\end{proof}

Note that $p^m_l \composed p^n_m = p^n_l$ whenever these maps are defined, and 
therefore $(\set{K_n}_n, \{p^n_m\}_{n - m > D + \delta})$ is an inverse system 
of simplicial complexes.

@@

In order to define projections from $\boundary G$ into our system $(K_n)$, we
first describe a sequence of open covers of $\boundary G$.

@@

\begin{definition}
  For each $n$, fix a partition of unity subordinate to the covering
  $\set{U_n(x)}_{x \in S_n}$ of $\boundary G$. This is gives a
  continuous map from $\boundary G$ to the nerve of the covering. Denote by
  $p^\infty_n\from \boundary G \to K_n$ the composition of this map with the
  inclusion of the nerve into $K_n$, as given by Lemma~\ref{lem:L_Ksimplicial}.
\end{definition}

@@

  Then let $y_1 = \gamma_1(n)$ and $y_2 = \gamma_2(n+1)$. We have
  $\dist(y_1, \gamma_2) \leq 2\delta$, so by the triangle inequality
  $\dist(y_1, y_2) \leq 4\delta +1$. It follows that $\dist(x_1, x_2)
  \leq 6\delta+3$, and so the claimed inequality holds by
  \cref{rem:dist_vs_newdist}.
\end{proof}

@@

  Let $p_1$ be a point on $[\alpha_1(n_1), \alpha_2(n_2)] \union \alpha_2$
  within a distance $\delta$ of $\alpha_1(m_1)$.  If $p_1$ is on $\alpha_2$
  then
  \begin{align*}
    \dist(\alpha_1(m_1), \alpha_2(m_2)) 
      & \leq \delta + \dist(p_1, \alpha_2(m_2)) \\
      & \leq 2\delta + \abs{m_1 - m_2}
  \end{align*}
  It follows that
  \begin{align*}
    \gromov{e}{\alpha_1(m_1)}{\alpha_2(m_2)} 
      &\geq \frac{1}{2}(m_1 + m_2 - 2\delta - \abs{m_1 - m_2})\\
      &\geq \min\{m_1, m_2\} - \delta
  \end{align*} 
  and the result follows. Otherwise $p_1$ is on $[\alpha_1(n_1),
  \alpha_2(n_2)]$. Similarly either the claimed result holds, or there is a
  point $p_2$ on $[\alpha_1(n_1), \alpha_2(n_2)]$ within a
  distance $\delta$ of $\alpha_1(m_2)$.

  If $p_1$ lies between $\alpha_2(n_2)$ and $p_2$ on $[\alpha_1(n_1),
  \alpha_2(n_2)]$ then, by considering the triangle with vertices $p_2$,
  $\alpha_2(m_2)$ and $\alpha_2(n_2)$, we see that $\dist(p_1, \alpha_2) \leq
  2\delta$. Therefore $\dist(\alpha_1(m_1), \alpha_2) \leq 3\delta$, and so
  \begin{equation*}
    \gromov{e}{\alpha_1(m_1)}{\alpha_2(m_2)} \geq \min\{m_1, m_2\} - 3\delta
  \end{equation*}
  and the result follows.

  Otherwise,
  \begin{equation*}
    \dist(\alpha_1(n_1), \alpha_2(n_2)) = \dist(\alpha_1(n_1), p_1) +
        \dist(p_1, p_2) + \dist(p_2, \alpha_2(n_2)).
  \end{equation*}
  Therefore,
  \begin{align*}
    \mathrlap{\gromov{e}{\alpha_1(n_1)}{\alpha_n(n_2)} - \gromov{e}{\alpha_1(m_1)}{\alpha_2(m_2)}}
    \quad\quad\phantom{{}={}}&\\
           &\mathllap{{}={}}\left(\dist(\alpha_1(m_1), \alpha_1(n_1)) - \dist(\alpha_1(m_1), p_1)\right) \\
           &+ \left(\dist(\alpha_1(m_1), \alpha_2(m_2)) - \dist(p_1, p_2)\right) \\
           &+ \left(\dist(\alpha_2(m_2), \alpha_2(n_2)) - \dist(p_2, \alpha_2(m_2))\right) \\
           &\mathllap{{}\leq{}} \delta+2\delta+\delta=4\delta 
  \end{align*}
  It follows that
  \begin{equation*}
    \gromov{e}{\alpha_1(\infty)}{\alpha_2(\infty)} \leq \gromov{e}{\alpha_1(m_1)}{\alpha_2(m_2)} + 6\delta,
  \end{equation*}
  which completes the proof.
\end{proof}

@@

  By definition of the visual metric, $\gromov{e}{\xi_1}{\xi_2} \geq
  \log_a(k_1/\rho_0)$. Therefore, by \cref{lem:boundary_gromov_product},
  \begin{equation*}
    \gromov{e}{\alpha_1(n)}{\alpha_2(n)} \geq \min\{n - 3\delta, \log_a(k_1/\rho_0) - 6\delta\}.
  \end{equation*}
  In particular,
  \begin{equation*}
    \dist(\alpha_1(n), \alpha_2(n)) \leq \max\{6\delta, 12\delta + 2(n - \log(k_1/\rho_0))\}
  \end{equation*}
  and the result follows.
\end{proof}

>>>>

\todo{Proof-read section 3 onwards}

@@

If a group acts geometrically on a proper hyperbolic space then, following
Gromov~\cite{gromov87}, we say that the group is hyperbolic, and quasi-isometry
invariants of the space can be thought of as invariants of the group. One
particularly powerful quasi-isometry invariant that can be used in
this way is the Gromov boundary of the space. This is the set of geodesic rays
in the space eminating from a fixed base point, where rays are defined to be
equivalent if they remain a bounded distance apart. In this way one can define
the boundary $\boundary G$ of a hyperbolic group $G$.

The boundary of a hyperbolic group is useful as a quasi-isometry invariant of
the group: if groups $G_1$ and $G_2$ are quasi-isometric then $\boundary G_1
\isom \boundary G_2$. Furthermore, certain topological properties of the
boundary translate directly into algebraic properties of the group. By a
theorem of Stallings~\cite{stallings68}, $\boundary G$ is disconnected if and
only if $G$ admits a splitting as an amalgamated product or HNN extension over
a finite subgroup. A theorem of Bowditch further characterises splittings over
virtually cyclic subgroups: if $\boundary G$ is connected then $G$ splits over
a virtually cyclic subgroup if and only if $\boundary G$ can be separated by
the removal of a pair of points. Sometimes the topology of $\boundary G$
determines $G$ even more precisely: if $\boundary G$ is homeomorphic to a
circle then the Convergence Group Theorem of Tukia, Casson, Jungreis and
Gabai~\cite{tukia88,cassonjungreis94,gabai92} tells us that $G$ contains a
surface group at finite index.

In~\cite{gromov87}, Gromov showed that the boundary of a hyperbolic group is a
compact metric space with finite topological dimension. Beyond these basic
properties, and excepting the case when $G$ is the fundamental group of a
negatively curved smooth manifold, in which case $\boundary G$ is a sphere,
boundaries of hyperbolic groups tend to be wild, fractal-like objects.
In~\cite{kapovichkleiner00}, Kapovich and Kleiner ask the following question.

\begin{question}\cite[Question A]{kapovichkleiner00}
  Which topological spaces are boundaries of hyperbolic groups?
\end{question}

On the one hand, finding concrete examples of spaces arising as boundaries has
proved difficult, but on the other, relatively few restrictions on the topology
of the boundary are known. One restriction of particular importance is a
theorem of Bestvina and Mess~\cite{bestvinamess91}, which states that if
$\boundary G$ is connected and is not disconnected by the removal of any single
point then $\boundary G$ is locally path connected. Combined with a result of
Bowditch~\cite{bowditch98b} and Swarup~\cite{swarup96}, this implies that
$\boundary G$ is locally path connected whenever it is connected.

The goal of this paper is to prove the following theorem, which restricts the
local complexity of the boundary of a hyperbolic group in terms of its global
complexity in a similar way to the theorem of Bestvina and Mess described
above.

@@

The paper is structured as follows. In Section~\ref{sec:spheres} we use the
Rips complex of the group $G$ to build a sequence $(K_n)_{n \geq 0}$ of
simplicial complexes that give successively better approximations to the
boundary $\boundary G$. These simplicial complexes are spheres in the Rips
complex. There exist natural simplicial maps $p^n_m\from K_n\to K_m$ for $n -
m$ large enough, and there exist natural continuous maps
$p^\infty_n\from\boundary G \to K_n$ for all $n$.  The remainder of
Section~\ref{sec:spheres} will be spent proving quantitative statements about
these maps.

In Section~\ref{sec:horoball} we prove a compactness result: given a sequence
of vertices $v_n \in K_n$, there exist group elements $g_n$ and a subcomplex
$H$ of the Rips complex such that translation by $g_n$ maps a large
neighbourhood in $K_n$ of $v_n$ onto a large neighbourhood in $H$ of $g_n\cdot
v_n$.  We may further assume that $(g_n)$ converges to a point $\xi \in
\boundary G$, and then we think of $H$ as a horoball in $G$ asymptotic to
$\xi$. We then spend the remainder of Section~\ref{sec:horoball} describing
projection maps $\boundary G - \set\xi \to H$. This relates the topology of
$\boundary G - \set\xi$ to that of $H$, and hence to the topology of large
balls in $K_n$ for $n$ sufficiently large.

In Section~\ref{sec:double_dagger} we begin by recalling the so-called
double-dagger condition $\ddag_M$ of Bestvina and Mess. This condition is
guaranteed to hold if $\boundary G$ does not contain a cut point, and implies
that $\boundary G$ is locally path connected. We then modify the condition to
allow for easier generalisation, and introduce a new condition $\S_M$, which we
will see to be related the local simple connectedness of $\boundary G$. Using
the compactness results from Section~\ref{sec:horoball} we show that this
condition is satisfied whenever we have that $\boundary G - \set\xi$ is simply
connected for any $\xi$ in $\boundary G$.

Then, in Section~\ref{sec:local_simple_connectedness} we show that our new
condition $\S_M$ implies the local simple connectedness of $\boundary G$. We
begin by using $\S_M$ to construct a sequence of maps $(i^n_\infty)_n$ from the
2-skeleton of $K_n$ to $\boundary G$. These maps can be thought of as
approximate sections of $p^\infty_n$, in that $i^n_\infty\composed p^\infty_n$
converges uniformly to the identity map $\boundary G \to \boundary G$ as $n$
tends to $\infty$. By controlling the diameter of the image under $i^n_\infty$
of a simplex in $K_n$ we are able to prove the local simple connectedness of
$\boundary G$.

@@

Let $G$ be a hyperbolic group and let $S$ be a symmetric generating set for
$G$. Let $e$ be the identity element of $G$, which we use as a base point. Let
$\dist(\cdot, \cdot)$ be the length metric on the Cayley graph of $G$ with
respect to $S$ obtained by setting all edges to have length $1$; on $G$ this
metric agrees with the word metric.

Let $\delta \geq 0$ be large enough to ensure that the Cayley graph of $G$ with
respect to $S$ is $\delta$-hyperbolic, meaning that all geodesic triangles in
the Cayley graph of $G$ are $\delta$-slim. (This means that each edge of each
triangle is contained in the $\delta$-neighbourhood of the union of the other
two edges.) Let $\boundary G$ be the boundary of the group, which can be
identified with the Gromov boundary of the Cayley graph of $G$. To reduce the
growth of constants later on we in fact assume that all \emph{ideal} triangles
(with vertices in $G \union \boundary G$) are $\delta$-hyperbolic; this can be
guaranteed by multiplying $\delta$ by $4$.

Given points $x$ and $y$ in $G \union \boundary G$ we will often denote by
$[x,y]$ a geodesic from $x$ to $y$, thought of as a subset of the Cayley graph
together with its boundary. Geodesics in hyperbolic spaces are not necessarily
unique, so this notation will stand for a choice of geodesic, but the choice
will not matter.

Throughout this and subsequent sections we use~\cite{bridsonhaefliger99} as a
reference for fundamental properties of hyperbolic spaces. The reader is
referred to~\cite{benaklikapovich00} for a useful survey on boundaries of
hyperbolic groups.

\begin{definition}
  Let $D = 10^6\delta + 10^6$.  We denote by $K$ the \emph{Rips complex} of
  $G$; this is the simplicial complex with vertex set $G$ such that a finite
  set of elements of $G$ spans a simplex if and only if that set has diameter
  at most $D$ with respect to the word metric $d$.
\end{definition}

@@

\begin{remark}\label{rem:dist_vs_newdist}
  The extended ``metric'' $\newdist$ is not a metric, since it only satisifes
  the triangle inequality up to a bounded additive error, but this will not
  cause a problem.

  Notice that $\newdist$ agrees with $\dist$ on $G$, and also that the choice
  of vertices $x^\star$ and $y\star$ in the definition does not matter very
  much: for any $x$ and $y$ in $K$, if $x^\star$ and $y^\star$ are an
  alternative choice of vertices of simplices containing $x$ and $y$
  respectively, then
  \begin{equation*}
    \abs{\newdist(x,y) - \dist(x^\star, y^\star)} \leq 2D
  \end{equation*}
\end{remark}

@@

\begin{definition}
  \todo{Extend to the boundary}
  Given points $x$, $y$ and $z$ in $K$, define the \emph{Gromov product
  $\gromov{x}{y}{z}$ of $x$, $y$ and $z$} as follows.
  \begin{equation*}
    \gromov{x}{y}{z} = \frac{1}{2}\left(\newdist(x,y) + \newdist(x,z) -
          \newdist(y,z)\right)
  \end{equation*}
\end{definition}

We now equip $K \union \boundary G$ with a topology as follows.

@@

\begin{definition}
  Fix an ordering on $S$. Let $S^\star$ be the free monoid on 
  $S$; order $S^\star$ by the lex-least ordering: this ordering is defined by 
  requiring that shorter words precede longer and that words of equal length are 
  ordered lexicographically.

  For $n \geq m$ define a map $p^n_m \from S_n \to S_m$ as follows.
  Given an element $g$ of $S_n$ let $w_g \in S^\star$ be the lex-least
  representative of $g$; note that this word in $S^\star$ necessarily has
  length $n$.  Then define $p^n_m(g)$ to be the element of $G$ given by
  truncating $w_g$ to a word of length $m$.
\end{definition}

@@

\begin{proof}
  Let $\set{g_1, \dotsc, g_k}$ span a simplex in $K_n$.  Then for $i$ and
  $j$ in $\set{1, \dotsc, k}$, the distance from $p^n_m(g_i)$ to any
  geodesic from $g_i$ to $g_j$ is at least $n - m - D$, which is greater than
  $\delta$, so by hyperbolicity of a triangle with vertices $e$, $g_i$ and $g_j$,
  the distance from $p^n_m(g_i)$ to any geodesic from $e$ to $g_j$ is at most
  $\delta$. We may choose this geodesic to contain $p^n_m(g_j)$, and $\dist(e,
  p^n_m(g_i)) = \dist(e, p^n_m(g_j))$, so $\dist(p^n_m(g_i), p^n_m(g_j)) \leq
  2\delta \leq D$.  It follows that the diameter of $\set{p^n_m(g_1), \dotsc,
  p^n_m(g_k)}$ is at most $D$ and therefore the set spans a simplex in $K_m$.
\end{proof}

\begin{remark}
  Note that $p^m_l \composed p^n_m = p^n_l$ whenever these maps are defined, and 
  therefore $(\set{K_n}_n, \{p^n_m\}_{n - m > D + \delta})$ is an inverse system 
  of simplicial complexes.
\end{remark}

@@

In order to define projections from $\boundary G$ into our system $(K_n)$, we
first describe a sequence of covers of $\boundary G$.

@@

\begin{definition}
  For each $n$, fix a partition of unity subordinate to the covering
  $\set{U_n(x)}_{x \in S_n}$ of $\boundary G$. Equivalently, this is a
  continuous map from $\boundary G$ to the nerve of the covering. Denote by
  $p^\infty_n\from \boundary G \to K_n$ the composition of this map with the
  inclusion of the nerve into $K_n$, as given by Lemma~\ref{lem:L_Ksimplicial}.
\end{definition}

@@

  Then let $y_1 = \gamma_1(n)$ and $y_2 = \gamma_2(n+1)$. By considering the
  degenerate triangle with edges $\gamma_1$ and $\gamma_2$,
  $\dist(y_1, \gamma_2) \leq \delta$, so by the triangle inequality
  $\dist(y_1, y_2) \leq 2\delta +1$. It follows that $\dist(x_1, x_2)
  \leq 6\delta+3$, and so the claimed inequality holds by
  \cref{rem:dist_vs_newdist}.
\end{proof}

@@

  Let $p_1$ be a point on $[\alpha_1(n_1), \alpha_2(n_2)] \union \alpha_2$
  within a distance $\delta$ of $\alpha_1(m_1)$.  If $p_1$ is on $\alpha_2$
  then
  \begin{align*}
    \dist(\alpha_1(m_1), \alpha_2(m_2)) 
      & \leq \delta + \dist(p_1, \alpha_2(m_2)) \\
      & \leq 2\delta + \abs{m_1 - m_2}
  \end{align*}
  It follows that
  \begin{align*}
    \gromov{e}{\alpha_1(m_1)}{\alpha_2(m_2)} 
      &\geq \frac{1}{2}(m_1 + m_2 - 2\delta - \abs{m_1 - m_2})\\
      &\geq \min\{m_1, m_2\} - \delta
  \end{align*} 
  and the result follows. Otherwise $p_1$ is on $[\alpha_1(n_1),
  \alpha_2(n_2)]$. Similarly either the claimed inequality holds, or there is a
  point $p_2$ on $[\alpha_1(n_1), \alpha_2(n_2)]$ within a distance $\delta$ of
  $\alpha_1(m_2)$.

  If $p_1$ lies between $\alpha_2(n_2)$ and $p_2$ on $[\alpha_1(n_1),
  \alpha_2(n_2)]$ then, by considering the triangle with vertices $p_2$,
  $\alpha_2(m_2)$ and $\alpha_2(n_2)$, we see that $\dist(p_1, \alpha_2) \leq
  2\delta$. Therefore $\dist(\alpha_1(m_1), \alpha_2) \leq 3\delta$, and so by
  a repetition of the argument used earlier,
  \begin{equation*}
    \gromov{e}{\alpha_1(m_1)}{\alpha_2(m_2)} \geq \min\{m_1, m_2\} - 3\delta
  \end{equation*}
  and the result follows.

  Otherwise,
  \begin{equation*}
    \dist(\alpha_1(n_1), \alpha_2(n_2)) = \dist(\alpha_1(n_1), p_1) +
        \dist(p_1, p_2) + \dist(p_2, \alpha_2(n_2)).
  \end{equation*}
  Therefore,
  \begin{align*}
    \mathrlap{\gromov{e}{\alpha_1(n_1)}{\alpha_n(n_2)} - \gromov{e}{\alpha_1(m_1)}{\alpha_2(m_2)}}
    \quad\quad\phantom{{}={}}&\\
           &\mathllap{{}={}}\left(\dist(\alpha_1(m_1), \alpha_1(n_1)) - \dist(\alpha_1(m_1), p_1)\right) \\
           &+ \left(\dist(\alpha_1(m_1), \alpha_2(m_2)) - \dist(p_1, p_2)\right) \\
           &+ \left(\dist(\alpha_2(m_2), \alpha_2(n_2)) - \dist(p_2, \alpha_2(m_2))\right) \\
           &\mathllap{{}\leq{}} \delta+2\delta+\delta=4\delta.
  \end{align*}
  It follows that
  \begin{equation*}
    \gromov{e}{\alpha_1(\infty)}{\alpha_2(\infty)} \leq \gromov{e}{\alpha_1(m_1)}{\alpha_2(m_2)} + 6\delta,
  \end{equation*}
  which completes the proof.
\end{proof}

@@

  By definition of the visual metric, $\gromov{e}{\xi_1}{\xi_2} \geq
  \log_a(k_1/\rho_0)$. Therefore, by \cref{lem:boundary_gromov_product},
  \begin{equation*}
    \gromov{e}{\alpha_1(n)}{\alpha_2(n)} \geq \min\{n - 3\delta, \log_a(k_1/\rho_0) - 6\delta\}.
  \end{equation*}
  In particular,
  \begin{equation*}
    \dist(\alpha_1(n), \alpha_2(n)) \leq \max\{6\delta, 12\delta + 2(n - \log(k_1/\rho_0))\}
  \end{equation*}
  and the result follows since $D \geq 12\delta$.
\end{proof}
